---
title: "analysis1"
author: "whalen"
date: "November 18, 2014"
output: html_document
---

Outliers have been replaced with `NA` (see `explore_report.Rmd` for details). Here I load the datasets with the `NA`s inserted and start setting things up for analyses - thin plate spline and/or regression kriging.

Load libraries I plan to use for data manipulation (`dplyr`) and plotting (`ggplot2`).
```{r load basic libraries, echo=TRUE}
library(ggplot2, suppressPackageStartupMessages(expr = TRUE))
library(plyr)
library(dplyr, suppressPackageStartupMessages(expr = TRUE))
```

Read-in the data sets, one daily and one hourly, and then plot.
```{r dataframe summaries, echo=TRUE}
# These data files have already had the outliers replaced with `NA`
dy_rg_data <- readRDS("dy_rg_data.rds")
summary(dy_rg_data)
qplot(id, daily_ppt, data=dy_rg_data) + theme_bw()

hr_rg_data <- readRDS("hr_rg_data.rds")
summary(hr_rg_data)
qplot(id, hourly_ppt, data=hr_rg_data)
```


Francesco talking over the exploration with me
```{r explore with francesco, eval=FALSE}
test=dy_rg_data$date2[order(dy_rg_data$date2), ]
test=dy_rg_data[order(dy_rg_data$date2), ]
head(test)
head(test,40)
test=dy_rg_data[order(dy_rg_data$date2,dy_rg_data$id), ]
```

A couple of rain gauges have been moved to new locations near their established areas over the years, so I have shapefiles corresponding the move of one or more gauge stations.
```{r list rain gague shapefiles}
list.files("shapefiles", pattern = "*.shx")
```
This shows that there are four shapefiles for various stretches of the study, 2003-2008, 2009-2011, 2012-2014, and then 2015 and beyond. The 2015 is the result of moving a rain gauge during our 2014 visits. I will not be using that file for the data I have in hand.


```{r breakdown years by shapefiles}
library(rgdal)# GDAL bindings for R; I believe this requires GDAL installed for your operating system
ogrListLayers("shapefiles")
rg_0308 <- readOGR(dsn="shapefiles", layer="RainGauges2003_2008")
str(rg_0308)
rg_0308@data$NAME
rg_0308@data$NAME <- tolower(rg_0308@data$NAME)
rg_0308@data$NAME[rg_0308@data$NAME=="macki"] <- "mackie"
rg_0308@data$NAME[rg_0308@data$NAME=="sdc"] <- "sec"
rg_0308@data$NAME[rg_0308@data$NAME=="sugar"] <- "sugarloaf"
unique(dy_rg_data$id)

month_rg_0308 <- dy_rg_data %>%
      group_by(id, year, month) %>% 
      filter(year<=2008) %>%
      summarise(monthly_ppt = sum(daily_ppt))
month_rg_0308$id <- as.factor(month_rg_0308$id)
summary(month_rg_0308)
```

```{r geostats}
library(sp)
library(raster)
library(rgdal)
elev <- raster("dem/sod15m.tif")
str(elev)
crs <- elev@crs
rg_0308 <- spTransform(rg_0308, CRS=crs)
str(rg_0308)
spplot(elev, scales=list(draw=T), sp.layout=list("sp.points", rg_0308))

rg_0308@data$elev <- extract(elev, rg_0308)
str(rg_0308)
#spplot(slope, scales=list(draw=T), sp.layout=list("sp.points", depth, pch="+"))
```

Joining monthly precipitation data to spatial points data frame
```{r joining data}
rg_0308@data <- inner_join(rg_0308@data, month_rg_0308, by = c("NAME" = "id"))
str(month_rg_0308)
str(rg_0308)
summary(rg_0308)
```

Plotting precipitation against elevation for a subset of the data
```{r plot elevation precip}
rg08 <- filter(rg_0308@data, year==2008)
qplot(rg_0308@data$elev, rg_0308@data$monthly_ppt)
qplot(elev, monthly_ppt, data = rg08, facets = .~ month, color = NAME, main = "2008 Rain")
qplot(elev, monthly_ppt, data = rg08 %>% filter(month==1), color = NAME, main = "January 2008 Rain") + geom_text(aes(label=NAME),hjust=-0.1, vjust=0)
```

Developing a linear modeling for kriging
```{r rk model jan 2008}
rg_0108 <- subset(rg_0308@data, year==2008 & month==1)
#coordinates(rg_0108) = ~NORTHING+EASTING
str(rg_0108)
rg_0308 <- readOGR(dsn="shapefiles", layer="RainGauges2003_2008")
rg_0308 <- spTransform(rg_0308, CRS=crs)
rg_0108@proj4string <- rg_0308@proj4string
rg_0108 <- spTransform(rg_0108, CRS=crs)

elev_ppt = overlay(as(elev, "SpatialGridDataFrame"), rg_0108)   # create grid-points overlay
   # copy the slope values
lm_rg_0108 <- lm(monthly_ppt ~ elev, rg_0108)
summary(lm_rg_0108)
```


```{r thin plate spline}
library(fields)
loc_matrix <- cbind(rg_0108$NORTHING,rg_0108$EASTING)
ppt <- c(rg_0108$monthly_ppt)
jan08tps <- Tps(loc_matrix, ppt)
summary(jan08tps)
plot(jan08tps)
surface(jan08tps)

look <- predict(jan08tps, lambda = 2.0)
look <- predict(jan08tps, df = 4)

tps1 <- Tps(loc_matrix, ppt, df = 4)
summary(tps1)
set.panel(2,2)
plot(tps1)
set.panel()

out_jan08 <- predictSurface(jan08tps)
image(out_jan08)

fit <- Tps(ozone$x, ozone$y)
set.panel(2,2)
plot(fit)

```
