---
title: "Rain gauge data exploration"
author: "whalen"
date: "November 10, 2014"
output: html_document
---

# Rain gauge data summary and manipulation
#### Goal: summarize precipitation events into hourly and daily, and convert to inches
- Conversion factor: 1 event = 0.01 inches


```{r Load libraries}
library(plyr)
library(data.table)
library(dplyr)
library(tidyr)
```

NOT RUN: Manipulations of some single files that were not correctly formatted for opening in Excel. I later found a better way to do this using the CSV library in Python that allows relatively simple conversion of text/CSV files from one delimiter to another.
```{r not run, echo=FALSE}
# d1 <- fread("Rain_Gauge/all_years_raw_exports/SDC_502836_2007_RG.csv")
# d1 <- separate(d1, date_time, c("date", "time"), sep=" ")
# write.csv(d1, "Rain_Gauge/all_years_raw_exports/SDC_502836_2007_RG.csv")
# 
# d1 <- fread("Rain_Gauge/all_years_raw_exports/SDC_502836_2008_RG.csv")
# d1$id <- "sec"
# write.csv(d1, "Rain_Gauge/all_years_raw_exports/SDC_502836_2008_RG.csv")
# 
# d1 <- fread("Rain_Gauge/all_years_raw_exports/SUGAR_20070416_20090411_RG.csv")
# d1$id <- "sugarloaf"
# d1 <- d1 %>% separate(date_time, c("date", "time"), sep=" ") %>% 
#       select(id, date, time, events)
# write.csv(d1, "Rain_Gauge/all_years_raw_exports/SUGAR_20070416_20090411_RG.csv")
# 
# d1 <- fread("Rain_Gauge/all_years_raw_exports/YAHNG_502848_2009_RG.txt")
# names(d1)[1:2]  <- c("dt", "events")
# d1$id <- "yahng"
# d1 <- d1 %>% separate(dt, c("date", "time"), sep=" ") %>% 
#       select(id, date, time, events)
# write.csv(d1, "Rain_Gauge/all_years_raw_exports/YAHNG_502848_2009_RG.csv")
# head(d1)
# 
# d1 <- fread("Rain_Gauge/all_years_raw_exports/MITSUI_502950_2006_RG.csv")
# head(d1)
# d1$id <- "mitsui"
# write.csv(d1, "Rain_Gauge/all_years_raw_exports/MITSUI_502950_2006_RG.csv")
```



