---
title: "Cross-validation of Precipitation Interpolation"
author: "Whalen Dillon"
date: "September 15, 2015"
output: html_document
---

Examine the outputs from the cross-validation procedure in `v.surf.rast` (2-dimensional interpolation) and `v.vol.rast` (3-dimensional interpolation) from GRASS-GIS. Like any cross-validation procedure, the goal is to estimate the predictive error for a given set of input parameters. I use scripts written by Jaro Hofierka and freely available on the [GRASS Book website](grassbook.org) to iterate through multiple combinations of tension and smoothing. While appearing as integer in these outputs, the smoothing value actually ranges from 0.1 to 0.9, i.e. a smoothing value of 10 equals 0.10.

In these cases, it is a leave-one-out cross-validation procedure, where each point in the dataset is excluded and the surface estimated from the remaining points. Then the difference between the actual and estimated value for each point is recorded. The root-mean square error and/or mean absolute error are two metrics that indicate how well the model performs in terms of accurate prediction using the given settings, with lower values indicating better accuracy.

I converted the original values in inches to millimeters, so the values reported here are in millimeter units.

# Cross-validation 
```{r load packages, echo=F, message=F}
library(plyr); library(dplyr); library(ggplot2)
```

```{r load 2D ppt amount files, echo=FALSE}
files <- list.files("cross_validation_grass/surf_rst/", full.names = T)
ppt2d_cv <- ldply(files, function(i){read.csv(i)})
```
```{r load 2D ppt days files, echo=F}
files <- list.files("cross_validation_grass/surf_rst_days/", full.names = T)
pptdys2d_cv <- ldply(files, function(i){read.csv(i)})
```

```{r remove dataframes, echo=F}
# rm(list = ls(pattern = "^rain"))
```

```{r load rain gauge shapefiles, echo=F, message=F}
library(rgdal)
shp_list <- list.files("shapefiles", pattern = "^[rain_]")
shp_list <- grep(".shp$", shp_list, value = TRUE)
shp_list <- shp_list[2:11]

shp_list <- strsplit(shp_list, ".shp")
tmp <- strsplit(unlist(shp_list), "rain_gauges")
tmp <- unlist(tmp)
tmp <- tmp[seq(2, 20, 2)]

n = 4
for(shp in shp_list){
      assign(paste("rainy", "0", n, sep = ""), 
             readOGR(dsn = "shapefiles", layer = shp, verbose = F))
      n = n + 1
}

rainy10 <- rainy010; rainy11 <- rainy011 
rainy12 <- rainy012; rainy14 <- rainy013
rm(list = ls(pattern = "01"))
rm(list = "tmp","shp","files","n","shp_list")
```

## Visualize cross-validation values from `v.surf.rast`
```{r plot cross validation of each rainy season, echo=F}
ggplot(ppt2d_cv, aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      facet_grid(year ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation Rainy Seasons Total Precipitation (v.surf.rst)") +
      theme_bw()

ggplot(pptdys2d_cv, aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      facet_grid(year ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation Rainy Seasons Days of Precipitation (v.surf.rst)") +
      theme_bw()
```

```{r min MAE 2D, echo=F}
# head(ppt2d_cv)
summary(ppt2d_cv)
ppt2d_cv$smoothing <- ppt2d_cv$smoothing/100

summary(pptdys2d_cv)
pptdys2d_cv$smoothing <- pptdys2d_cv$smoothing/100

ppt2d_cvmin <- rbind(
      filter(ppt2d_cv, year == 2004) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2005) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2006) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2007) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2008) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2009) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2010) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2011) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2012) %>% filter(mean_abs == min(mean_abs)),
      filter(ppt2d_cv, year == 2014) %>% filter(mean_abs == min(mean_abs))
)
knitr::kable(ppt2d_cvmin, caption = "Minimum MAE of precipitation estimation for each rainy season")

pptdys2d_cvmin <- rbind(
      filter(pptdys2d_cv, year == 2004) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2005) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2006) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2007) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2008) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2009) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2010) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2011) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2012) %>% filter(mean_abs == min(mean_abs)),
      filter(pptdys2d_cv, year == 2014) %>% filter(mean_abs == min(mean_abs))
)
knitr::kable(pptdys2d_cvmin, caption = "Minimum MAE of rainy days for each rainy season")
```


This gives the values for the minimum MAE (mean_abs) for each year from the cross-validation and the associated parameters. Since there is a lot of variability in the MAE especially for tension values between 1 and 10, I am going to get parameters minimum MAE at each tension value. This would result in 24 surfaces to qualitatively examine for each season, however, it appears that in many cases the MAE value no longer changes much after a certain tension is reached. If I can roughly identify this "break" then I can reduce the number of surfaces for qualitative comparison. (*I think this will likely be the same for assessing the results from 3-D interpolations*)


### 2004 2D Cross-Validation
```{r plot summary 2004 2D cross-validation, echo=F}
##### 2004 2-D Cross-validation
ggplot(filter(ppt2d_cv, year == 2004), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2004 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2004))

ggplot(filter(pptdys2d_cv, year == 2004), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2004 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2004))
```

Based on these graphs I am going to examine the minimum and maximum MAE for tension values 1-10 and 20-90 for total precipitation and number of days of precipitation.

```{r 2004 minimum MAE for selected tension values, echo=F}
## Total precipitation 
# minimum MAE for tension 1-10
mae.min.2004 <- rbind(filter(ppt2d_cv, year == 2004, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2004, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2004), min_mae)

knitr::kable(mae.min.2004, digits = 2, caption = "Minimum MAE Values - 2004 total precipitation")

## Days of precipitation
# minimum MAE for tension 1-10
mae.dys.min.2004 <- rbind(filter(pptdys2d_cv, year == 2004, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2004, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2004), min_mae)

knitr::kable(mae.min.2004, digits = 2, caption = "Minimum MAE Values - 2004 days")

# # tension = 2 minimum MAE parameters
# filter(ppt2d_cv, year == 2004, tension == 2) %>% filter(mean_abs == min(mean_abs))
# # mae=164.11 t=2 s=0.2
# 
# # tension = 3 minimum MAE parameters
# filter(ppt2d_cv, year == 2004, tension == 3) %>% filter(mean_abs == min(mean_abs))
# # mae=161.68 t=3 s=0.5
# 
# # tension = 4 MAE parameters
# filter(ppt2d_cv, year == 2004) %>% filter(between(mean_abs, 183.5, 183.6))
# # mae=183.56 t=50 s=0.8; mae=183.52 t=7 s=0.5
# 
# # find ~3rd Qu. MAE parameters
# filter(ppt2d_cv, year == 2004) %>% filter(between(mean_abs, 186.15, 186.5))
# # mae=186.28 t=40 s=0.4; mae=186.39 t=30 s=0.9
# 
# # maximum MAE parameters
# filter(ppt2d_cv, year == 2004) %>% filter(mean_abs == max(mean_abs))
# # mae=268.79 t=10 s=10
# filter(ppt2d_cv, year == 2004, tension > 10) %>% filter(mean_abs == max(mean_abs))
# # mae=208.8 t=20 s=0.1
```

```{r 2004 max MAE for tension values, echo=F}
## Total precipitation
mae.max.2004 <- rbind(filter(ppt2d_cv, year == 2004, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2004, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2004), max_mae)

knitr::kable(mae.max.2004, digits = 2, caption = "Maximum MAE Values - 2004")
```


### 2005 2D Cross-Validation
```{r plot summary 2005 2D cross validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2005), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2005 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2005))

ggplot(filter(pptdys2d_cv, year == 2005), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2005 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2005))
```

For this 2005 data I am going to include tension values out to 100 for total precipitation and out to 70 for rainy days. (It just crossed my mind to examine surfaces with maximum MAE for a given value of tension to provide a contrast of the effect from the smoothing value)

```{r 2005 minimum MAE for tension values, echo=F}
## Total precipitation
mae.min.2005 <- rbind(filter(ppt2d_cv, year == 2005, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2005, between(tension, 20, 100)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2005), min_mae)

knitr::kable(mae.min.2005, digits = 2, caption = "Minimum MAE Values - 2005 total precipitation")

## Days of precipitation
mae.dys.min.2005 <- rbind(filter(pptdys2d_cv, year == 2005, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2005, between(tension, 20, 70)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2005), min_mae)

knitr::kable(mae.dys.min.2005, digits = 2, caption = "Minimum MAE Values - 2005 - rainy days")


# # minimum MAE parameters
# filter(ppt2d_cv, year == 2005) %>% filter(mean_abs == min(mean_abs))
# # mae=171.80 t=1 s=0.9
# filter(ppt2d_cv, year == 2005, tension >= 10) %>% filter(mean_abs == min(mean_abs))
# # mae=184.8 t=150 s=0.9
# 
# # mean MAE parameters
# filter(ppt2d_cv, year == 2005) %>% filter(between(mean_abs, 198, 199))
# # mae=198.36 t=5 s=0.9; mae=198.37 t=30 s=0.7
# 
# # 1st Qu. MAE parameters
# filter(ppt2d_cv, year == 2005, between(mean_abs, 187, 188))
# # mae=187.7 t=90 s=0.9
```
```{r 2005 maximum MAE for tension values, echo=F}
## Total precipitation
mae.max.2005 <- rbind(filter(ppt2d_cv, year == 2005, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,100,10]
filter(ppt2d_cv, year == 2005, between(tension, 20, 100)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2005), max_mae)

knitr::kable(mae.max.2005, digits = 2, caption = "Maximum MAE Values - 2005")
```


### 2006 2-D Cross-validation
```{r plot summary 2006 2D cross validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2006), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2006 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2006))

ggplot(filter(pptdys2d_cv, year == 2006), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2006 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2006))
```

I am selecting tension values 1-10 and 20-60 for total precipitation, and out to 90 for rainy days

```{r 2006 minimum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.min.2006 <- rbind(filter(ppt2d_cv, year == 2006, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,60,10]
filter(ppt2d_cv, year == 2006, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2006), min_mae)

knitr::kable(mae.min.2006, digits = 2, caption = "Minimum MAE Values - 2006 total precipitation")

## Rainy days
mae.dys.min.2006 <- rbind(filter(pptdys2d_cv, year == 2006, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2006, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2006), min_mae)

knitr::kable(mae.dys.min.2006, digits = 2, caption = "Minimum MAE Values - 2006 rainy days")


# # minimum MAE parameters
# filter(ppt2d_cv, year == 2006) %>% filter(mean_abs == min(mean_abs))
# # mae=243.47 t=1 s=0.1
# filter(ppt2d_cv, year == 2006, tension >= 10) %>% filter(mean_abs == min(mean_abs))
# # mae=291.08 t=90 s=0.9
# 
# # 1st Qu. MAE parameters
# filter(ppt2d_cv, year == 2006, between(mean_abs, 291.08, 291.1))
# # mae=291.1 t=80 s=0.9
# 
# filter(ppt2d_cv, year == 2006, tension == 50)
# # mae=291.75 t=50 s=0.9; mae=292.48 t=50 s=0.5
```
```{r 2006 maximum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.max.2006 <- rbind(filter(ppt2d_cv, year == 2006, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,60,10]
filter(ppt2d_cv, year == 2006, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2006), max_mae)

knitr::kable(mae.max.2006, digits = 2, caption = "Maximum MAE Values - 2006 total precipitation")

## Rainy days
mae.dys.max.2006 <- rbind(filter(pptdys2d_cv, year == 2006, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2006, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.dys.max.2006), max_mae)

knitr::kable(mae.dys.max.2006, digits = 2, caption = "Maximum MAE Values - 2006 rainy days")
```


### 2007 2-D Cross-validation
```{r plot summary 2007 2D cross-validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2007), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2007 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2007))

ggplot(filter(pptdys2d_cv, year == 2007), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2007 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2007))
```

For total precipitation the MAE values start to cluster more strongly around a tension value of 5; I am going to summarize tension values 1-10 and 20-100. Clustering of MAE values for rainy days doesn't start until tension values of 20, but tighten up more quickly. I am going to assess tension values out to 90.

```{r 2007 minimum MAE parameters for select tension values}
## Total precipitation
mae.min.2007 <- rbind(filter(ppt2d_cv, year == 2007, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,100,10]
filter(ppt2d_cv, year == 2007, between(tension, 20, 100)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2007), min_mae)

knitr::kable(mae.min.2007, digits = 2, caption = "Minimum MAE Values - 2007 total precipitation")

## Rainy days
mae.dys.min.2007 <- rbind(filter(pptdys2d_cv, year == 2007, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2007, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2007), min_mae)

knitr::kable(mae.dys.min.2007, digits = 2, caption = "Minimum MAE Values - 2007 rainy days")

# # minimum MAE parameters
# filter(ppt2d_cv, year == 2007) %>% filter(mean_abs == min(mean_abs))
# # mae=96.69 t=1 s=0.9
# filter(ppt2d_cv, year == 2007, tension >= 10) %>% filter(mean_abs == min(mean_abs))
# # mae=102.38 t=150 s=0.9
# 
# # 1st Qu. MAE parameters
# filter(ppt2d_cv, year == 2007, between(mean_abs, 102.9, 103.2))
# # mae=103.12 t=90 s=0.9
# 
# # mean MAE parameters
# filter(ppt2d_cv, year == 2007, between(mean_abs, 105.4, 105.6))
# # mae=105.5 t=30 s=0.9; mae=105.54 t=40 s=0.3
```
```{r 2007 maximum MAE parameters for select tension values}
## Total rainfall
mae.max.2007 <- rbind(filter(ppt2d_cv, year == 2007, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2007, between(tension, 20, 100)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2007), max_mae)

knitr::kable(mae.max.2007, digits = 2, caption = "Maximum MAE Values - 2007")

## Rainy days
mae.dys.max.2007 <- rbind(filter(pptdys2d_cv, year == 2007, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2007, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.dys.max.2007), max_mae)

knitr::kable(mae.dys.max.2007, digits = 2, caption = "Maximum MAE Values - 2007 rainy days")
```


### 2008 2-D Cross-validation
```{r plot summary 2008 cross-validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2008), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2008 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2008))

ggplot(filter(pptdys2d_cv, year == 2008), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2008 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2008))
```

Selecting tension values 1-10 and 20-80 for total precipitation, and out to 90 for rainy days.

```{r 2008 minimum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.min.2008 <- rbind(filter(ppt2d_cv, year == 2008, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,80,10]
filter(ppt2d_cv, year == 2008, between(tension, 20, 80)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2008), min_mae)

knitr::kable(mae.min.2008, digits = 2, caption = "Minimum MAE Values - 2008 total precipitation")

# filter(ppt2d_cv, year == 2008) %>% filter(mean_abs == min(mean_abs))
# filter(ppt2d_cv, year == 2008, tension >= 9) %>% filter(mean_abs == min(mean_abs))

## Rainy days
mae.dys.min.2008 <- rbind(filter(pptdys2d_cv, year == 2008, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2008, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2008), min_mae)

knitr::kable(mae.dys.min.2008, digits = 2, caption = "Minimum MAE Values - 2008 rainy days")
```
```{r 2008 maximum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.max.2008 <- rbind(filter(ppt2d_cv, year == 2008, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,80,10]
filter(ppt2d_cv, year == 2008, between(tension, 20, 80)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2008), max_mae)

knitr::kable(mae.max.2008, digits = 2, caption = "Maximum MAE Values - 2008 total precipitation")

## Rainy days
mae.max.2008 <- rbind(filter(pptdys2d_cv, year == 2008, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2008, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2008), max_mae)

knitr::kable(mae.max.2008, digits = 2, caption = "Maximum MAE Values - 2008 rainy days")
```


### 2009 2-D Cross-validation
```{r plot summary 2009 2d cross-validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2009), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2009 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2009))

ggplot(filter(pptdys2d_cv, year == 2009), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2009 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2009))
```

This is the reverse of the pattern previously seen, where overall MAE has decreased with increasing tension for total precipitation. I'm going to select values 1-10 and 20-70 for total precipitation. 

For rainy days, MAE generally decreases with increasing tension, specifically for tension values from 20-150. I'm going to examine tension values 1-10 and 90-150 for rain days estimation.

```{r 2009 minimum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.min.2009 <- rbind(filter(ppt2d_cv, year == 2009, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,70,10]
filter(ppt2d_cv, year == 2009, between(tension, 20, 70)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2009), min_mae)

knitr::kable(mae.min.2009, digits = 2, caption = "Minimum MAE Values - 2009 total precipitation")
# 
# filter(ppt2d_cv, year == 2009) %>% filter(mean_abs == min(mean_abs))
# filter(ppt2d_cv, year == 2009, tension >= 10) %>% filter(mean_abs == min(mean_abs))

## Rainy days
mae.dys.min.2009 <- rbind(filter(pptdys2d_cv, year == 2009, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,70,10]
filter(pptdys2d_cv, year == 2009, between(tension, 90, 150)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2009), min_mae)

knitr::kable(mae.dys.min.2009, digits = 2, caption = "Minimum MAE Values - 2009 rainy days")
```
```{r 2009 maximum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.max.2009 <- rbind(filter(ppt2d_cv, year == 2009, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2009, between(tension, 20, 70)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2009), max_mae)

knitr::kable(mae.max.2009, digits = 2, caption = "maximum MAE Values - 2009 total precipitation")

## Rainy days
mae.dys.max.2009 <- rbind(filter(pptdys2d_cv, year == 2009, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2009, between(tension, 90, 150)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2009), max_mae)

knitr::kable(mae.max.2009, digits = 2, caption = "maximum MAE Values - 2009 rainy days")
```


### 2010 2-D Cross-validation
```{r plot summary 2010 cross-validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2010), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2010 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2010))

ggplot(filter(pptdys2d_cv, year == 2010), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2010 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(pptdys2d_cv, year == 2010))
```

This has a similar pattern to the 2009 data, though all the MAE values for tension 1-10 are greater than for all other values of tension. I am going to examine tension values 1-10 and 20-90 for total precipitation.

Rainy days has a pattern more like earlier years. I will examine tension values 1-10 and 20-100.

```{r 2010 minimum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.min.2010 <- rbind(filter(ppt2d_cv, year == 2010, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2010, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2010), min_mae)

knitr::kable(mae.min.2010, digits = 2, caption = "Minimum MAE Values - 2010 total precipitation")
# 
# filter(ppt2d_cv, year == 2010) %>% filter(mean_abs == min(mean_abs))
# filter(ppt2d_cv, year == 2010, tension < 20) %>% filter(mean_abs == min(mean_abs))

## Rainy days
mae.dys.min.2010 <- rbind(filter(pptdys2d_cv, year == 2010, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,100,10]
filter(pptdys2d_cv, year == 2010, between(tension, 20, 100)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2010), min_mae)

knitr::kable(mae.min.2010, digits = 2, caption = "Minimum MAE Values - 2010 rainy days")
```
```{r 2010 maximum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.max.2010 <- rbind(filter(ppt2d_cv, year == 2010, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2010, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2010), max_mae)

knitr::kable(mae.max.2010, digits = 2, caption = "Maximum MAE Values - 2010 total precipitation")

## Rainy days
mae.dys.max.2010 <- rbind(filter(pptdys2d_cv, year == 2010, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(pptdys2d_cv, year == 2010, between(tension, 20, 90)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2010), max_mae)

knitr::kable(mae.max.2010, digits = 2, caption = "Maximum MAE Values - 2010 rainy days")
```


### 2011 2-D Cross-validation
```{r plot summary 2011 2D cross-validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2011), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2011 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2011))

ggplot(filter(pptdys2d_cv, year == 2011), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2011 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2011))
```

Again, a similar pattern to 2009 and 2010 data. The clustering gets very tight for smoothing values as soon as tension values reach 20. I am going to examine tension values 1-10 and 20-60.

For rainy days I am going to examine tension values 1-10 and 20-50.

```{r 2011 minimum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.min.2011 <- rbind(filter(ppt2d_cv, year == 2011, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,60,10]
filter(ppt2d_cv, year == 2011, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2011), min_mae)

knitr::kable(mae.min.2011, digits = 2, caption = "Minimum MAE Values - 2011 total precipitation")

## Rainy days
mae.dys.min.2011 <- rbind(filter(pptdys2d_cv, year == 2011, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,50,10]
filter(pptdys2d_cv, year == 2011, between(tension, 20, 50)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2011), min_mae)

knitr::kable(mae.dys.min.2011, digits = 2, caption = "Minimum MAE Values - 2011 rainy days")
# 
# filter(ppt2d_cv, year == 2011) %>% filter(mean_abs == min(mean_abs))
# filter(ppt2d_cv, year == 2011, tension >= 10) %>% filter(mean_abs == min(mean_abs))
```
```{r 2011 maximum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.max.2011 <- rbind(filter(ppt2d_cv, year == 2011, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,60,10]
filter(ppt2d_cv, year == 2011, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2011), max_mae)

knitr::kable(mae.max.2011, digits = 2, caption = "Maximum MAE Values - 2011 total precipitation")

## Rainy days
mae.dys.max.2011 <- rbind(filter(pptdys2d_cv, year == 2011, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,50,10]
filter(pptdys2d_cv, year == 2011, between(tension, 20, 50)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.dys.max.2011), max_mae)

knitr::kable(mae.dys.max.2011, digits = 2, caption = "Maximum MAE Values - 2011 rainy days")
```


### 2012 2-D Cross-validation
```{r plot summary 2012 2D cross-validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2012), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2012 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2012))

ggplot(filter(pptdys2d_cv, year == 2012), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2012 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(pptdys2d_cv, year == 2012))
```

Some interesting patterns going on with these data. For tension values 1-10 there is a U-shape for values of MAE as values of tension increase, particularly for smoothing values > 0.5. Overall, the pattern is similar to 2009, 2010, and 2011 where MAE values increase with increasing tension values. I am going to examing tension values 1-10 and 20-60.

For rainy days I am examining tension values 1-10 and 20-60 as well.

```{r 2012 minimum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.min.2012 <- rbind(filter(ppt2d_cv, year == 2012, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,60,10]
filter(ppt2d_cv, year == 2012, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2012), min_mae)

knitr::kable(mae.min.2012, digits = 2, caption = "Minimum MAE Values - 2012 total precipitation")

## Rainy days
mae.dys.min.2012 <- rbind(filter(pptdys2d_cv, year == 2012, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,60,10]
filter(pptdys2d_cv, year == 2012, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2012), min_mae)

knitr::kable(mae.dys.min.2012, digits = 2, caption = "Minimum MAE Values - 2012 rainy days")
# filter(ppt2d_cv, year == 2012) %>% filter(mean_abs == min(mean_abs))
# filter(ppt2d_cv, year == 2012, tension < 30) %>% filter(mean_abs == min(mean_abs))
```
```{r 2012 maximum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.max.2012 <- rbind(filter(ppt2d_cv, year == 2012, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,60,10]
filter(ppt2d_cv, year == 2012, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2012), max_mae)

knitr::kable(mae.max.2012, digits = 2, caption = "Maximum MAE Values - 2012 total precipitation")

## Rainy days
mae.dys.max.2012 <- rbind(filter(pptdys2d_cv, year == 2012, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,60,10]
filter(pptdys2d_cv, year == 2012, between(tension, 20, 60)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2012), max_mae)

knitr::kable(mae.dys.max.2012, digits = 2, caption = "Maximum MAE Values - 2012 rainy days")
```


### 2014 2-D Cross-validation
```{r plot summary 2014 2D cross-validation, echo=F}
ggplot(filter(ppt2d_cv, year == 2014), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2014 Rainy Season Total Precipitation (v.surf.rst)") +
      theme_bw()
summary(filter(ppt2d_cv, year == 2014))

ggplot(filter(pptdys2d_cv, year == 2014), aes(x = tension, y = mean_abs)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2014 Rainy Season Rainy Days (v.surf.rst)") +
      theme_bw()
summary(filter(pptdys2d_cv, year == 2014))
```

Similar overall pattern to 2009-2012 data with increasing MAE values as values for tension increase for total rainfall. I'm going to examine tension values 1-10 and 20-70.

For rainy days, MAE generally decreases with increasing tension. I'm going to examine tension values 1-10 and 70-150.

```{r 2014 minimum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.min.2014 <- rbind(filter(ppt2d_cv, year == 2014, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [20,70,10]
filter(ppt2d_cv, year == 2014, between(tension, 20, 70)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.min.2014), min_mae)

knitr::kable(mae.min.2014, digits = 2, caption = "Minimum 2D MAE Values - 2014 total precipitation")
# filter(ppt2d_cv, year == 2014) %>% filter(mean_abs == min(mean_abs))

## Rainy days
mae.dys.min.2014 <- rbind(filter(pptdys2d_cv, year == 2014, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs)),
# minimum MAE for tension [70,150,10]
filter(pptdys2d_cv, year == 2014, between(tension, 70, 150)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))
) %>% rename(min_mae = mean_abs)
arrange(ungroup(mae.dys.min.2014), min_mae)

knitr::kable(mae.dys.min.2014, digits = 2, caption = "Minimum 2D MAE Values - 2014 rainy days")
```
```{r 2014 maximum MAE parameters for select tension values, echo=F}
## Total precipitation
mae.max.2014 <- rbind(filter(ppt2d_cv, year == 2014, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [20,90,10]
filter(ppt2d_cv, year == 2014, between(tension, 20, 70)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.max.2014), max_mae)

knitr::kable(mae.max.2014, digits = 2, caption = "Maximum 2D MAE Values - 2014 total precipitation")

## Rainy days
mae.dys.max.2014 <- rbind(filter(pptdys2d_cv, year == 2014, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs)),
# maximum MAE for tension [70,150,10]
filter(pptdys2d_cv, year == 2014, between(tension, 70, 150)) %>% group_by(tension) %>% filter(mean_abs == max(mean_abs))
) %>% rename(max_mae = mean_abs)
arrange(ungroup(mae.dys.max.2014), max_mae)

knitr::kable(mae.dys.max.2014, digits = 2, caption = "Maximum 2D MAE Values - 2014 rainy days")
```


## Visualize cross-validation values from `v.vol.rast`
```{r load 3D ppt amount files, echo=FALSE}
files <- list.files("cross_validation_grass/vol_zrst/", full.names = T)
ppt3d_cvz <- ldply(files, function(i){read.csv(i)})
```
```{r load 3D ppt days files, echo=F}
files <- list.files("cross_validation_grass/vol_zrst_days/", full.names = T)
pptdys3d_cvz <- ldply(files, function(i){read.csv(i)})
rm(files)
```

```{r summarize 3D, echo=F}
# head(ppt3d_cvz)
summary(ppt3d_cvz)
ppt3d_cvz$smoothing <- ppt3d_cvz$smoothing/100
ppt3d_cvz <- rename(ppt3d_cvz, mae = mean_abs, zscale = scale)

# Create dataframe of each year's minimum MAE
ppt3d_cvmin <- rbind(
      filter(ppt3d_cvz, year == 2004) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2005) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2006) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2007) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2008) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2009) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2010) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2011) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2012) %>% filter(mae == min(mae)),
      filter(ppt3d_cvz, year == 2014) %>% filter(mae == min(mae))
)
ppt3d_cvmin

summary(pptdys3d_cvz)
pptdys3d_cvz$smoothing <- pptdys3d_cvz$smoothing/100
pptdys3d_cvz <- rename(pptdys3d_cvz, mae = mean_abs, zscale = scale)

# Create dataframe of each year's minimum MAE
pptdys3d_cvmin <- rbind(
      filter(pptdys3d_cvz, year == 2004) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2005) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2006) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2007) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2008) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2009) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2010) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2011) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2012) %>% filter(mae == min(mae)),
      filter(pptdys3d_cvz, year == 2014) %>% filter(mae == min(mae))
)
pptdys3d_cvmin
```


### Cross-validation v.vol.rst: 2004 total precipitation
```{r plot summary 3d total ppt 2004 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2004), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2004 Rainy Season Total Precipitation (v.vol.rst)") +
      theme_bw()
# summary(filter(ppt3d_cv, year == 2004))

ggplot(filter(ppt3d_cvz, year == 2004), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      # facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2004 Rainy Season (v.vol.rst)") +
      theme_bw()

ggplot(filter(ppt3d_cvz, year == 2004), aes(x = tension, y = mae)) +
      geom_point(aes(color = (smoothing))) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2004 Rainy Season (v.vol.rst)") +
      theme_bw()
# MAE in terms of smoothing seems to be all over the place, indicating effect of zscale(?)

ggplot(filter(ppt3d_cvz, year == 2004, tension <= 10), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 10, 1)) +
      ggtitle("Cross-validation 2004 Rainy Season (v.vol.rst) for tension <= 10") +
      theme_bw()
```

Small values for the zscale parameter appear to consistently provide the lowest values for MAE across values of tension and smoothing, except for an increase with tension values from 1 to ~5 when the smoothing parameter was set to 0.1. Higher values of the smoothing parameter appears to stabilize the relationship with zscale, where increasing zscale increases MAE. 

For values of tension between 1 and 10 the MAE values are higher for higher zscale values across all values of tension and smoothing. Smoothing appears to have a strong effect on the base minimum MAE, where MAE for a given value of tension and zscale appears to increase with the value of the smoothing parameter. This seems to be true most clearly for smoothing values > 0.2. Over all the tension values, MAE shows a consistent increase with increasing tension.

According to this I want fairly low values for tension, zscale, and smoothing parameters, but I'm not at all sure what the surface will look like (i.e. does it represent a realistic/acceptable spatial pattern of precipitation?)

I'm going to start with looking at tension values between 1 and 10, because the 10 lowest values for MAE fall into this range. This will give me parameters for 10 surfaces to compare.

```{r 2004 3d MAE minimums for select values of tension, echo=F}
# filter(ppt2d_cv, year == 2014, between(tension, 1, 10)) %>% group_by(tension) %>% filter(mean_abs == min(mean_abs))

mae3d.min.2004 <- filter(ppt3d_cvz, year == 2004, between(tension, 1, 10)) %>% 
      group_by(tension) %>% filter(mae == min(mae)) %>% 
      rename(min_mae = mae)
mae3d.min.2004 <- ungroup(mae3d.min.2004)

knitr::kable(arrange(mae3d.min.2004, min_mae), caption = "Minimum 3D MAE Values - 2004 total precipitation")
# 
# summary(mae3d.min.2004)
# filter(mae3d.min.2004, min_mae == max(min_mae))

# filter(ppt3d_cv, year == 2004, tension == 1)
# filter(ppt3d_cv, year == 2004, tension == 2) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cv, year == 2004, tension == 3) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cv, year == 2004, tension == 4) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cv, year == 2004, tension == 5) %>% filter(mean_abs == min(mean_abs))
```

The 10 values for minimum MAE for each of the tension values are all smaller than the 1st quartile of MAE for all of the 2004 cross-validation data, which is `r 198`.

### Cross-validation v.vol.rst: 2004 rainy days
```{r plot summary 3d days ppt 2004 CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2004), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2004 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
# summary(filter(ppt3d_cv, year == 2004))
```

For the rainy days cross validation MAE decreases with increasing tension and smoothing values, while simultaneously being lower for small values of zscale (< 25). I'm going to get parameters of MAE for tension values 1-150, which appear to be all for a smoothing value of 0.9.

```{r 2004 rainy days 3d MAE minimums for select values of tension, echo=F}
mae3d.dys.min.2004 <- filter(pptdys3d_cvz, year == 2004, between(tension, 1, 150)) %>% 
      group_by(tension) %>% filter(mae == min(mae)) %>% 
      rename(min_mae = mae)
mae3d.dys.min.2004 <- ungroup(mae3d.dys.min.2004)

knitr::kable(arrange(mae3d.dys.min.2004, min_mae), caption = "Minimum 3D MAE Values - 2004 rainy days")
```


### Cross-validation v.vol.rst: 2005 total precipitation
```{r plot 3d 2005 total ppt CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2005), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2005 Rainy Season Total Precipitation (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2005))
```

Increasing the value of smoothing decreases MAE. Generally, higher values for tensions also decreases MAE, but this is not consistent for tension values between 1 and 10. For these tension values, MAE increases up to about 3 or 4, then begins to decline. The smallest values for MAE don't appear to vary dramtically for tension values from 20 to 150, though they do get smaller.

```{r 2005 3d MAE total ppt minimums for select values of tension, echo=F}
head(filter(ppt3d_cvz, year == 2005) %>% arrange(mae), 10)
mae3d.min.2005 <- filter(ppt3d_cvz, year == 2005, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2005, caption = "Minimum 3D MAE Values - 2005 total rainfall")
# filter(ppt3d_cvz, year == 2005, tension > 3) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2005, tension < 3) %>% filter(mean_abs == min(mean_abs))
```

### Cross-validation v.vol.rst: 2005 rainy days
```{r plot 3d 2005 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2005), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2005 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2005))
```

For the rainy days, MAE generally decreases with increasing tension values from 1-6 or 7 when the smoothing parameter is 0.4 or greater, however, MAE begins to increase again as tension gets larger. Lower values of zscale correspond with lower values of MAE across all combinations of tension and smoothing. The smallest MAE value corresponds to tension = 10, smoothing = 0.1, and zscale = 1.

```{r 2005 3d MAE days minimums for select values of tension, echo=F}
# head(filter(pptdys3d_cvz, year == 2005) %>% arrange(mae), 10)
mae3d.dys.min.2005 <- filter(pptdys3d_cvz, year == 2005, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.dys.min.2005, caption = "Minimum 3D MAE Values - 2005 rainy days")
```

The range of smallest MAE across all tension values is only about 0.7, i.e., less than one day. The majority are when smoothing = 0.1.

### Cross-validation v.vol.rst: 2006 total precipitation
```{r plot 3d 2006 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2006), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2006 Rainy Season (v.vol.rst)") +
      theme_bw()
# summary(filter(ppt3d_cvz, year == 2006))
```

Interestingly the relationship between zscale and MAE becomes somewhat inverted as the values for tension increase. The lowest values of zscale correspond with the higher values of MAE, but it appears that a middle value results in the lowest value of MAE when tension is > 3 This pattern appears consistent across values of smoothing. All of the lowest values for MAE are for tension values <= 10.

```{r 2006 3d MAE minimums, echo=F}
mae3d.min.2006 <- filter(ppt3d_cvz, year == 2006, between(tension, 1, 10)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2006, caption = "Minimum 3D MAE Values - 2006")

# filter(ppt3d_cvz, year == 2006, tension <= 10) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2006, tension == 10) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2006, tension == 20) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2006, tension == 30) %>% filter(mean_abs == min(mean_abs))
```

### Cross-validation v.vol.rst: 2006 rainy days
```{r plot 3d 2006 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2006), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2006 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2006))
```

Higher values of smoothing result in smaller values of MAE. Generally smaller values of zscale also correspond with smaller values of MAE. MAE initially increases with increasing tension up to tension = 4, but tension = 1 results in one of the lowest values of MAE.

```{r 2006 3d MAE days minimums for select values of tension, echo=F}
# head(filter(pptdys3d_cvz, year == 2005) %>% arrange(mae), 10)
mae3d.dys.min.2006 <- filter(pptdys3d_cvz, year == 2006, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2006, caption = "Minimum 3D MAE Values - 2006 rainy days")
```

The maximum range across all tension values is a little less than one day, about 0.9.

### Cross-validation v.vol.rst: 2007 total precipitation
```{r plot 3d 2007 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2007), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2007 Rainy Season (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2007))
```

```{r 2007 3d MAE minimums, echo=F}
mae3d.min.2007 <- filter(ppt3d_cvz, year == 2007, between(tension, 1, 20)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2007, caption = "Minimum 3D MAE Values - 2007")

# filter(ppt3d_cvz, year == 2007) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2007, tension ) %>% filter(mean_abs == min(mean_abs))
# produced only "overshoot" results; >100mm off of the given measurements
```

### Cross-validation v.vol.rst: 2007 rainy days
```{r plot 3d 2007 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2007), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2007 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2007))
```

MAE decreseases with increasing tension values from 2 to about 20, while tension values of one result in an MAE value less than when tension has values from 2-4 depending on the corresponding value for smoothing. Higher values of smoothing shrinks the variation in MAE values and generally results in smaller values of MAE.

```{r 2007 3d MAE days minimums for select values of tension, echo=F}
# head(filter(pptdys3d_cvz, year == 2005) %>% arrange(mae), 10)
mae3d.dys.min.2007 <- filter(pptdys3d_cvz, year == 2007, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2007, caption = "Minimum 3D MAE Values - 2007 rainy days")
```

Variability of minimum MAE values across all tension parameters is less than 0.4 days. Most of the minimum values are when smoothing = 0.1.

### Cross-validation v.vol.rst: 2008 total precipitation
```{r plot 3d 2008 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2008), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (mm)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2008 Rainy Season (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2008))
```

```{r 2008 3d MAE minimums, echo=F}
mae3d.min.2008 <- filter(ppt3d_cvz, year == 2008, between(tension, 1, 20)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2008, caption = "Minimum 3D MAE Values - 2008")

# filter(ppt3d_cvz, year == 2008) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2008) %>% filter(mean_abs < 5.513)
```

### Cross-validation v.vol.rst: 2008 rainy days
```{r plot 3d 2008 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2008), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2008 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2008))
```

Generally MAE decreases with increasing tension, with a nonlinear pattern for tension values 1-10. Higher values of smoothing also result in smaller values of MAE.

```{r 2008 3d MAE days minimums for select values of tension, echo=F}
mae3d.dys.min.2008 <- filter(pptdys3d_cvz, year == 2008, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2008, caption = "Minimum 3D MAE Values - 2008 rainy days")
```

Variability of minimum MAE across all values of tension is about 1.4 days. All of the minimum values occur when smoothing = 0.9.

### Cross-validation v.vol.rst: 2009 total precipitation
```{r plot 3d 2009 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2009), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error") +
      ggtitle("2009 Rainy Season (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2009))
```

```{r 2009 3d MAE minimums, echo=F}
mae3d.min.2009 <- filter(ppt3d_cvz, year == 2009, between(tension, 1, 20)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2009, caption = "Minimum 3D MAE Values - 2009")

# filter(ppt3d_cvz, year == 2009) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2009) %>% filter(mean_abs < 9.56)
```

### Cross-validation v.vol.rst: 2009 rainy days
```{r plot 3d 2009 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2009), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2009 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2009))
```

Relationships between MAE and other parameters are pretty variable. MAE decreases with tension values from 1 to 20 or 30 depending on the value of the smoothing parameter. The relationship with the zscale parameter switches with increasing values of tension, where for low tension values (< 7-10) a low zscale parameter results in lower MAE, and as tension increases (> 20 or 30) higher values of zscale result in lower MAE.

```{r 2009 3d MAE days minimums for select values of tension, echo=F}
mae3d.dys.min.2009 <- filter(pptdys3d_cvz, year == 2009, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2009, caption = "Minimum 3D MAE Values - 2009 rainy days")
```

Variability of minimum MAE values across all tension parameters is about 1.2 days. The zscale parameter is mostly >= 60, while 0.1 is the most frequent value for the smoothing parameter.

### Cross-validation v.vol.rst: 2010 total precipitation
```{r plot 3d 2010 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2010), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error") +
      ggtitle("2010 Rainy Season (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2010))
```

```{r 2010 3d MAE minimums, echo=F}
mae3d.min.2010 <- filter(ppt3d_cvz, year == 2010, between(tension, 1, 30)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2010, caption = "Minimum 3D MAE Values - 2010")
# filter(ppt3d_cvz, year == 2010) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2010) %>% filter(mean_abs < 6.594)
```

### Cross-validation v.vol.rst: 2010 rainy days
```{r plot 3d 2010 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2010), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2010 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2010))
```

Increasing values for smoothing result in smaller values of MAE, with a very narrow spread of MAE values when smoothing = 0.9. There is nonolinearity in MAE values with tension-zscale combinations when tension values are < 10. 

```{r 2010 3d MAE days minimums for select values of tension, echo=F}
mae3d.dys.min.2010 <- filter(pptdys3d_cvz, year == 2010, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2010, caption = "Minimum 3D MAE Values - 2010 rainy days")
```

The variability of minimum MAE values across values for tension is < 0.3 days. The majority of zscale values are >= 100, while 0.1 and 0.9 are the most common smoothing values.

### Cross-validation v.vol.rst: 2011 total precipitation
```{r plot 3d 2011 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2011), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error") +
      ggtitle("2011 Rainy Season (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2011))
```


```{r 2011 3d MAE minimums, echo=F}
mae3d.min.2011 <- filter(ppt3d_cvz, year == 2011, between(tension, 1, 30)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2011, caption = "Minimum 3D MAE Values - 2011")

# filter(ppt3d_cvz, year == 2011) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2011) %>% filter(mean_abs < 12.85)
```

### Cross-validation v.vol.rst: 2011 rainy days
```{r plot 3d 2011 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2011), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2011 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2011))
```

Generally higher values of all three parameters result in lower MAE. 

```{r 2011 3d MAE days minimums for select values of tension, echo=F}
mae3d.dys.min.2011 <- filter(pptdys3d_cvz, year == 2011, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2011, caption = "Minimum 3D MAE Values - 2011 rainy days")
```

Variability of minimum MAE values across all tension parameter values is about 0.6 days. All zscale values are >= 20 and all smoothing values are either 0.1 or 0.9, with 0.9 being the most frequent.

### Cross-validation v.vol.rst: 2012 total precipitation
```{r plot 3d 2012 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2012), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error") +
      ggtitle("2012 Rainy Season (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2012))
```


```{r 2012 3d MAE minimums, echo=F}
mae3d.min.2012 <- filter(ppt3d_cvz, year == 2012, between(tension, 1, 10)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2012, caption = "Minimum 3D MAE Values - 2012")

# filter(ppt3d_cvz, year == 2012) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2012) %>% filter(mean_abs < 4.932)
```

### Cross-validation v.vol.rst: 2012 rainy days
```{r plot 3d 2012 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2012), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2012 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2012))
```

As tension increases from 1-10 MAE decreases. Higher values of smoothing and zscale also tends to reduce MAE. 

```{r 2012 3d MAE days minimums for select values of tension, echo=F}
mae3d.dys.min.2012 <- filter(pptdys3d_cvz, year == 2012, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2011, caption = "Minimum 3D MAE Values - 2012 rainy days")
```

Variability in minimum MAE across tension values is about 0.6 days, smoothing parameter values are all either 0.1 or 0.9, and zscale values range between 20 and 90.

### Cross-validation v.vol.rst: 2014 total precipitation
```{r plot 3d 2014 CV, echo=F}
ggplot(filter(ppt3d_cvz, year == 2014), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error") +
      ggtitle("2014 Rainy Season (v.vol.rst)") +
      theme_bw()
summary(filter(ppt3d_cvz, year == 2014))
```


```{r 2014 3d MAE minimums, echo=F}
mae3d.min.2014 <- filter(ppt3d_cvz, year == 2014, between(tension, 1, 10)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)

knitr::kable(mae3d.min.2014, caption = "Minimum 3D MAE Values - 2014")

# filter(ppt3d_cvz, year == 2014) %>% filter(mean_abs == min(mean_abs))
# filter(ppt3d_cvz, year == 2014) %>% filter(mean_abs < 8.735)
```

### Cross-validation v.vol.rst: 2014 rainy days
```{r plot 3d 2014 rainy days CV, echo=F}
ggplot(filter(pptdys3d_cvz, year == 2014), aes(x = tension, y = mae)) +
      geom_point(aes(color = (zscale))) +
      facet_grid(smoothing ~.) +
      ylab("Mean Absolute Error (days)") +
      scale_x_continuous(breaks = seq(0, 150, 10)) +
      ggtitle("Cross-validation 2014 Rainy Season Rainy Days (v.vol.rst)") +
      theme_bw()
summary(filter(pptdys3d_cvz, year == 2014))
```

Values of MAE generally decrease with increasing tension and smoothing. Lower values of zscale (< 20) mostly appear to produce smaller values of MAE.

```{r 2014 3d MAE days minimums for select values of tension, echo=F}
mae3d.dys.min.2014 <- filter(pptdys3d_cvz, year == 2014, between(tension, 1, 150)) %>%
      group_by(tension) %>% filter(mae == min(mae)) %>%
      ungroup(.) %>% 
      arrange(mae)
knitr::kable(mae3d.dys.min.2014, caption = "Minimum 3D MAE Values - 2014 rainy days")
```

Variability of minimum MAE values is about 0.8 days, all smoothing values are 0.9, and nearly all of the zscale values are 1 with two exceptions, one of 40 and one of 60. The two outlying zscale values correspond to tension values of 1 and 2 respectively.

## PRISM Comparison to Measured Values
```{r prism comparison, echo=F, eval=F}
summary(rainy04)
plot(totppt_mm ~ psm_totppt, data = rainy04@data, main = 2004)
cor(rainy04@data$totppt_mm, rainy04@data$psm_totppt)

plot(totppt_mm ~ psm_totppt, data = rainy05@data, main = 2005)
cor(rainy05@data$totppt_mm, rainy05@data$psm_totppt)

plot(totppt_mm ~ psm_totppt, data = rainy06@data, main = 2006)
cor(rainy06@data$totppt_mm, rainy06@data$psm_totppt)

plot(totppt_mm ~ psm_totppt, data = rainy07@data, main = 2007)
cor(rainy07@data$totppt_mm, rainy07@data$psm_totppt)

plot(totppt_mm ~ psm_totppt, data = rainy08@data, main = 2008)
cor(rainy08@data$totppt_mm, rainy08@data$psm_totppt)

plot(totppt_mm ~ psm_totppt, data = rainy09@data, main = 2009)
cor(rainy09@data$totppt_mm, rainy09@data$psm_totppt)

plot(totppt_mm ~ psm_totppt, data = rainy10@data, main = 2010)
cor(rainy10@data$totppt_mm, rainy10@data$psm_totppt)
```

