---
title: "spatial_analysis_report"
author: "Whalen Dillon"
date: "October 2, 2015"
output: html_document
---

## Report to summarize some of the spatial analysis
This reports on some of the exploratory analysis of spatial relationships in the precipitation data, as well as a comparison of our rain gauge data to estimates from downscaled PRISM data. 
(These code chunks are mostly pulled from the `spatial_analysis.Rmd`)

```{r load spatial .Rdata, echo=FALSE}
load("~/Documents/soco_ppt/spatial_201510.RData")
```

#### Comparing Rainy Season Totals from PRISM & Rain Gauges
```{r rainy season correlation}
rs_totals.cor <- ggplot(data = rs_totals, aes(x = psm_totppt, y = totppt_mm))
rs_totals.cor + 
      geom_point() +
      geom_smooth(method ="lm", size = 0.5, se = F) +
      geom_abline(intercept = 0) +
      facet_grid(year ~ ., scales = "free_x", space = "free_x") +
      # scale_x_continuous(breaks = seq(500, 2000, 500)) +
      geom_text(aes(x = 500, y = 1000, label = label), data = correlations, 
                hjust = 0, vjust = 0) +
      xlab("PRISM precipitation (mm)") +
      ylab("Rain gauge precipitation (mm)") +
      ggtitle("Rainy Season Rain Gauge Precipitation vs. PRISM Precipiation") +
      theme_bw()
```

Table of rainy season aggregations correlations and deviations 
```{r rainy season cor table}
library(xtable)
xtable(correlations)
```

#### Comparing Monthly Totals from PRISM & Rain Gauges
```{r monthly correlation}
monthly.cor <- ggplot(data = prism_rg, aes(x = prism_ppt, y = monthly_ppt_mm))
monthly.cor +
      geom_point() +
      geom_smooth(method ="lm", size = 0.5, se = F) +
      geom_abline(intercept = 0) +
      facet_grid(month ~ year) +
      geom_text(aes(x = 200, y = 500, label = label), data = monthly.cor) +
      xlab("PRISM Precipitation (mm)") +
      ylab("Rain gauge precipitation (mm)") +
      ggtitle("Monthly Totals Rain Gauges vs. PRISM Precipitation") +
      theme_bw()
```

```{r}
monthly.cor
```

